---
title: "Assessment 2 - Report"
author: "Colin W"
date: "2022-08-15"
output: 
  bookdown::html_document2:
    code_folding: hide
---

```{r setup, R.options=knitr::opts_chunk$set(warnings=FALSE, message=FALSE, fig.width=12, fig.height=8)}
```



```{r Packages}
#load packages
library(tidyverse)
library(viridis)
library(plotly)
library(bookdown)

#load data
RawData <- read.csv("Data/Dataset3_Assessment3.csv")
```


## Visualising the shape of the data
Data is firstly plotted into a density plot to check whether it is normally distributed to be able to calculate summary statistics correctly.  

```{r fig1, fig.cap = 'density plot showing distribution of force (N) during a maximal adductor strength test', fig.height = 3, fig.width = 8, fig.align = "center"}
#create density plot
ggplot(data = RawData, aes(x = Force)) +
  geom_density() +
  labs(y = " ") +
  theme_light()
```
In Figure \@ref(fig:fig1) we can see that the data is in a bell curve shape. This means we will be able to accurately calculate summary stats as we have a uniform data set.  
&nbsp;  
&nbsp;  

## Creating graphs to provide context to data

Now that the data has been shown to be normally distributed we can start plotting it into graphs that will provide us with more information.
```{r fig2, fig.cap='Interactive box and whisker plot seperated by limb allowing comparison between participants and summary statistics', fig.height = 5, fig.width = 8, fig.align = "center"}
#create an overlaid box and whisker plot
ForcePlot <- ggplot(RawData, aes(x = Measure, y = Force)) +
  geom_jitter(aes(colour = Name)) +
  labs(x = " ") +
  scale_colour_viridis_d() +
  geom_boxplot(alpha = 0.3) +
  labs(x = " ") +
  theme_light() +
  facet_wrap(~Limb)

#make an interactive plot
ggplotly(ForcePlot)
```
&nbsp;  
In Figure \@ref(fig:fig2) we are able to highlight specific names in our dataset to compare them to other participants. We can compare them between legs as well as comparing them to summary statistics such as the median and upper and lower quartile. We can see that there is minimal data outside the upper and lower fences meaning there are few outliers within our dataset, which backs up what we see in Figure \@ref(fig:fig1).
&nbsp;  

## Data change over time

&nbsp;  
The data can be plotted by date to provide information about how the participants output changed over time.
```{r fig3, fig.cap= 'line plot showing seperate limb data for Kaitlin and Victoria', fig.height = 5, fig.width = 10, fig.align = "center"}
#convert date format
RawData$Date2 <- as.Date(RawData$Date, tryFormats = c("%d/%m/%Y"))

#create Dataset of 2 athletes to compare
KaitlinVictoria <- RawData %>% 
  filter(Name == "Kaitlin" | Name == "Victoria")
#create graph of 2 athletes
ggplot(data = KaitlinVictoria, aes(x = Date2, y = Force, colour = Name)) +
geom_point() +
geom_line() +
scale_colour_viridis_d() +
  facet_wrap(~Limb) +
theme_light() 
```
&nbsp;  
In the above Figure \@ref(fig:fig3) I have chosen to compare the data of Kaitlin and Victoria to see what observations can be made about how their output changed over time in respect to eachother.

* There is a break in the data between the 16th of December and the 13th of January, after this there appears to be a decrease in the output of both participants.

* Before the break the data for both participants overall is on a downward trend. This may be due to fatigue due to a high workload leading up to the break for instance in a pre-season sports setting.

* Overall Kaitlin's output is higher than Victoria's.

* Data across both of Kaitlin's limbs is quite similar in shape, however there is a clear strength variation with her right leg having a higher output overall. 

* Data across both of Victoria's limbs is also quite similar in shape apart from the 3rd and 4th datapoints after the break. There is less of a variation between legs for Kaitlin indicating even strength between both legs.

&nbsp; 

## Packages used 
Tidyverse - used to create plots within project via ggplot  
Viridis - used to create inclusive plots for people with colourblindness  
Plotly - used to to create interactive plots to better isolate data  
Bookdown - used to reference plots within text.
```{r}
citation(package = "base", lib.loc = NULL, auto = NULL)
citation(package = "tidyverse", lib.loc = NULL, auto = NULL)
citation(package = "viridis", lib.loc = NULL, auto = NULL)
citation(package = "plotly", lib.loc = NULL, auto = NULL)
citation(package = "bookdown", lib.loc = NULL, auto = NULL)
```

